Pulmonary acini exhibit complex changes during postnatal rat lung development

### 2022-01-25 manuscript.tex
Add CITATION.bib [ci skip]
This build is based on
https://github.com/habi/acinar-analysis-manuscript/commit/fc05b91d5caaddc9e4ccf93ef536f8f962a20170.

This commit was created by the following CI build and job:
https://github.com/habi/acinar-analysis-manuscript/commit/fc05b91d5caaddc9e4ccf93ef536f8f962a20170/checks
https://github.com/habi/acinar-analysis-manuscript/runs/1745825667

This manuscript
(\href{https://habi.github.io/acinar-analysis-manuscript/v/f1c96b127bb47e033309ae3205b2ddf5f0210582/}{permalink})
was automatically generated
from \href{https://github.com/habi/acinar-analysis-manuscript/tree/f1c96b127bb47e033309ae3205b2ddf5f0210582}{habi/acinar-analysis-manuscript@f1c96b1}
on November 10, 2021.

@@

\begin{itemize}
\item
  \textbf{David Haberthür}
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-3388-9187}{0000-0003-3388-9187}
  · \href{mailto:haberthuer@ana.unibe.ch?subject=Feedback\%20regarding\%20the\%20acinus\%20manuscript\%20(f1c96b1)}{haberthuer@ana.unibe.ch}
  · \includegraphics{images/github.svg}
  \href{https://github.com/habi}{habi}
  · \includegraphics{images/twitter.svg}
  \href{https://twitter.com/habi}{habi}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Eveline Yao}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Sébastien F. Barré}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Tiziana P. Cremona}
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-0061-9578}{0000-0003-0061-9578}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Stefan A. Tschanz}
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-3880-4437}{0000-0003-3880-4437}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Johannes C. Schittny}✉
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-4025-3961}{0000-0003-4025-3961}
  · \href{mailto:schittny@ana.unibe.ch?subject=Feedback\%20regarding\%20the\%20acinus\%20manuscript\%20(f1c96b1)}{schittny@ana.unibe.ch}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\end{itemize}

@@

\leavevmode\hypertarget{ref-ODQCVC23}{}%
2. Schittny JC. How high resolution 3-dimensional imaging changes our understanding of postnatal lung development. Histochemistry and Cell Biology. 2018;150: 677--691. doi:\href{https://doi.org/10.1007/s00418-018-1749-7}{10.1007/s00418-018-1749-7}

\leavevmode\hypertarget{ref-auNiE56G}{}%
3. Burri PH, Dbaly J, Weibel ER. The postnatal growth of the rat lung. I. Morphometry. The Anatomical Record. 1974;178: 711--730. doi:\href{https://doi.org/10.1002/ar.1091780405}{10.1002/ar.1091780405}

\leavevmode\hypertarget{ref-1AhvLCPky}{}%
4. Kauffman SL, Burri PH, Weibel ER. The postnatal growth of the rat lung II. Autoradiography. The Anatomical Record. 1974;180: 63--76. doi:\href{https://doi.org/10.1002/ar.1091800108}{10.1002/ar.1091800108}

\leavevmode\hypertarget{ref-XQLPrPTJ}{}%
5. Schittny JC. Development of the lung. Cell and Tissue Research. 2017;367: 427--444. doi:\href{https://doi.org/10.1007/s00441-016-2545-0}{10.1007/s00441-016-2545-0}

@@

\leavevmode\hypertarget{ref-BbEFwEVD}{}%
8. Haefeli-Bleuer B, Weibel ER. Morphometry of the human pulmonary acinus. The Anatomical Record. 1988;220: 401--414. doi:\href{https://doi.org/10.1002/ar.1092200410}{10.1002/ar.1092200410}

@@

\leavevmode\hypertarget{ref-uFNlWogb}{}%
11. Barré SF, Haberthür D, Stampanoni M, Schittny JC. Efficient estimation of the total number of acini in adult rat lung. Physiological Reports. 2014;2: e12063. doi:\href{https://doi.org/10.14814/phy2.12063}{10.14814/phy2.12063}

\leavevmode\hypertarget{ref-CVl41LwO}{}%
12. Fishler R, Hofemeier P, Etzion Y, Dubowski Y, Sznitman J. Particle dynamics and deposition in true-scale pulmonary acinar models. Scientific Reports. 2015;5: 14071. doi:\href{https://doi.org/10.1038/srep14071}{10.1038/srep14071}

\leavevmode\hypertarget{ref-eb0gk6VO}{}%
13. Hasler D, Anagnostopoulou P, Nyilas S, Latzin P, Schittny J, Obrist D. A multi-scale model of gas transport in the lung to study heterogeneous lung ventilation during the multiple-breath washout test. PLOS Computational Biology. 2019;15: e1007079. doi:\href{https://doi.org/10.1371/journal.pcbi.1007079}{10.1371/journal.pcbi.1007079}

@@

\leavevmode\hypertarget{ref-CPhMy3Ie}{}%
18. Woodward JD, Maina JN. A 3D digital reconstruction of the components of the gas exchange tissue of the lung of the muscovy duck, Cairina moschata. Journal of Anatomy. 2005;206: 477--492. doi:\href{https://doi.org/10.1111/j.1469-7580.2005.00413.x}{10.1111/j.1469-7580.2005.00413.x}

\leavevmode\hypertarget{ref-1GJECnKBu}{}%
19. Rodriguez M, Bur S, Favre A, Weibel ER. Pulmonary acinus: Geometry and morphometry of the peripheral airway system in rat and rabbit. American Journal of Anatomy. 1987;180: 143--155. doi:\href{https://doi.org/10.1002/aja.1001800204}{10.1002/aja.1001800204}

@@

\leavevmode\hypertarget{ref-RGBeCf8v}{}%
22. Kizhakke Puliyakote AS, Vasilescu DM, Newell JD, Wang G, Weibel ER, Hoffman EA. Morphometric differences between central vs. surface acini in A/J mice using high-resolution micro-computed tomography. Journal of Applied Physiology. 2016;121: 115--122. doi:\href{https://doi.org/10.1152/japplphysiol.00317.2016}{10.1152/japplphysiol.00317.2016}

@@

\leavevmode\hypertarget{ref-knMAkPPz}{}%
27. Marone F, Hintermüller C, McDonald S, Abela R, Mikuljan G, Isenegger A, et al. X-ray Tomographic Microscopy at TOMCAT. Journal of Physics: Conference Series. 2009;186: 012042. doi:\href{https://doi.org/10.1088/1742-6596/186/1/012042}{10.1088/1742-6596/186/1/012042}

@@

\leavevmode\hypertarget{ref-VELl7OiR}{}%
29. Haberthür D, Hintermüller C, Marone F, Schittny JC, Stampanoni M. Radiation dose optimized lateral expansion of the field of view in synchrotron radiation X-ray tomographic microscopy. Journal of Synchrotron Radiation. 2010;17: 590--599. doi:\href{https://doi.org/10.1107/s0909049510019618}{10.1107/s0909049510019618}

\leavevmode\hypertarget{ref-dNc8FfNn}{}%
30. Hsia CCW, Hyde DM, Ochs M, Weibel ER. An Official Research Policy Statement of the American Thoracic Society/European Respiratory Society: Standards for Quantitative Assessment of Lung Structure. American Journal of Respiratory and Critical Care Medicine. 2010;181: 394--418. doi:\href{https://doi.org/10.1164/rccm.200809-1522st}{10.1164/rccm.200809-1522st}

@@

\leavevmode\hypertarget{ref-nPoQ2EIB}{}%
33. TSCHANZ SA, BURRI PH, WEIBEL ER. A simple tool for stereological assessment of digital images: the STEPanizer. Journal of Microscopy. 2011;243: 47--59. doi:\href{https://doi.org/10.1111/j.1365-2818.2010.03481.x}{10.1111/j.1365-2818.2010.03481.x}

\leavevmode\hypertarget{ref-QiAxY2i3}{}%
34. Hyde DM, Tyler NK, Putney LF, Singh P, Gundersen HJG. Total number and mean size of alveoli in mammalian lung estimated using fractionator sampling and unbiased estimates of the Euler characteristic of alveolar openings. The Anatomical Record. 2004;277A: 216--226. doi:\href{https://doi.org/10.1002/ar.a.20012}{10.1002/ar.a.20012}

@@

\leavevmode\hypertarget{ref-FE9HLB4f}{}%
36. CRUZ-ORIVE LM. Precision of Cavalieri sections and slices with local errors. Journal of Microscopy. 1999;193: 182--198. doi:\href{https://doi.org/10.1046/j.1365-2818.1999.00460.x}{10.1046/j.1365-2818.1999.00460.x}

@@

\leavevmode\hypertarget{ref-IkHrgIj3}{}%
41. SHAPIRO SS, WILK MB. An analysis of variance test for normality (complete samples). Biometrika. 1965;52: 591--611. doi:\href{https://doi.org/10.1093/biomet/52.3-4.591}{10.1093/biomet/52.3-4.591}

\leavevmode\hypertarget{ref-MyzxBJ57}{}%
42. Mann HB, Whitney DR. On a Test of Whether one of Two Random Variables is Stochastically Larger than the Other. The Annals of Mathematical Statistics. 1947;18: 50--60. doi:\href{https://doi.org/10.1214/aoms/1177730491}{10.1214/aoms/1177730491}

@@

\leavevmode\hypertarget{ref-8Miti2Gz}{}%
44. Virtanen P, Gommers R, Oliphant TE, Haberland M, Reddy T, Cournapeau D, et al. SciPy 1.0: fundamental algorithms for scientific computing in Python. Nature Methods. 2020;17: 261--272. doi:\href{https://doi.org/10.1038/s41592-019-0686-2}{10.1038/s41592-019-0686-2}

@@

\leavevmode\hypertarget{ref-I9TmP6IU}{}%
46. Zeltner TB, Bertacchini M, Messerli A, Burri PH. Morphometric Estimation of Regional Differences in the Rat Lung. Experimental Lung Research. 2009;16: 145--158. doi:\href{https://doi.org/10.3109/01902149009087879}{10.3109/01902149009087879}

\leavevmode\hypertarget{ref-Zv1XcgAO}{}%
47. Borisova E, Lovric G, Miettinen A, Fardin L, Bayat S, Larsson A, et al. Micrometer-resolution X-ray tomographic full-volume reconstruction of an intact post-mortem juvenile rat lung. Histochemistry and Cell Biology. 2020;155: 215--226. doi:\href{https://doi.org/10.1007/s00418-020-01868-8}{10.1007/s00418-020-01868-8}

@@

\leavevmode\hypertarget{ref-eioib1TQ}{}%
49. Sznitman J, Sutter R, Altorfer D, Stampanoni M, Rösgen T, Schittny JC. Visualization of respiratory flows from 3D reconstructed alveolar airspaces using X-ray tomographic microscopy. Journal of Visualization. 2010;13: 337--345. doi:\href{https://doi.org/10.1007/s12650-010-0043-0}{10.1007/s12650-010-0043-0}

@@

\leavevmode\hypertarget{ref-YuJbg3zO}{}%
59. Himmelstein DS, Rubinetti V, Slochower DR, Hu D, Malladi VS, Greene CS, et al. Open collaborative writing with Manubot. PLOS Computational Biology. 2019;15: e1007128. doi:\href{https://doi.org/10.1371/journal.pcbi.1007128}{10.1371/journal.pcbi.1007128}
\end{cslreferences}

>>>>

This manuscript
(\href{https://habi.github.io/acinar-analysis-manuscript/v/fc05b91d5caaddc9e4ccf93ef536f8f962a20170/}{permalink})
was automatically generated
from \href{https://github.com/habi/acinar-analysis-manuscript/tree/fc05b91d5caaddc9e4ccf93ef536f8f962a20170}{habi/acinar-analysis-manuscript@fc05b91}
on January 25, 2022.

@@

\begin{itemize}
\item
  \textbf{David Haberthür}
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-3388-9187}{0000-0003-3388-9187}
  · \href{mailto:haberthuer@ana.unibe.ch?subject=Feedback\%20regarding\%20the\%20acinus\%20manuscript\%20(fc05b91)}{haberthuer@ana.unibe.ch}
  · \includegraphics{images/github.svg}
  \href{https://github.com/habi}{habi}
  · \includegraphics{images/twitter.svg}
  \href{https://twitter.com/habi}{habi}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Eveline Yao}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Sébastien F. Barré}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Tiziana P. Cremona}
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-0061-9578}{0000-0003-0061-9578}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Stefan A. Tschanz}
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-3880-4437}{0000-0003-3880-4437}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Johannes C. Schittny}✉
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-4025-3961}{0000-0003-4025-3961}
  · \href{mailto:schittny@ana.unibe.ch?subject=Feedback\%20regarding\%20the\%20acinus\%20manuscript\%20(fc05b91)}{schittny@ana.unibe.ch}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\end{itemize}

@@

\leavevmode\hypertarget{ref-ODQCVC23}{}%
2. Schittny JC. How high resolution 3-dimensional imaging changes our understanding of postnatal lung development. Histochem Cell Biol. 2018;150: 677--691. doi:\href{https://doi.org/10.1007/s00418-018-1749-7}{10.1007/s00418-018-1749-7}

\leavevmode\hypertarget{ref-auNiE56G}{}%
3. Burri PH, Dbaly J, Weibel ER. The postnatal growth of the rat lung. I. Morphometry. Anat Rec. 1974;178: 711--730. doi:\href{https://doi.org/10.1002/ar.1091780405}{10.1002/ar.1091780405}

\leavevmode\hypertarget{ref-1AhvLCPky}{}%
4. Kauffman SL, Burri PH, Weibel ER. The postnatal growth of the rat lung II. Autoradiography. Anat Rec. 1974;180: 63--76. doi:\href{https://doi.org/10.1002/ar.1091800108}{10.1002/ar.1091800108}

\leavevmode\hypertarget{ref-XQLPrPTJ}{}%
5. Schittny JC. Development of the lung. Cell Tissue Res. 2017;367: 427--444. doi:\href{https://doi.org/10.1007/s00441-016-2545-0}{10.1007/s00441-016-2545-0}

@@

\leavevmode\hypertarget{ref-BbEFwEVD}{}%
8. Haefeli-Bleuer B, Weibel ER. Morphometry of the human pulmonary acinus. Anat Rec. 1988;220: 401--414. doi:\href{https://doi.org/10.1002/ar.1092200410}{10.1002/ar.1092200410}

@@

\leavevmode\hypertarget{ref-uFNlWogb}{}%
11. Barré SF, Haberthür D, Stampanoni M, Schittny JC. Efficient estimation of the total number of acini in adult rat lung. Physiol Rep. 2014;2: e12063. doi:\href{https://doi.org/10.14814/phy2.12063}{10.14814/phy2.12063}

\leavevmode\hypertarget{ref-CVl41LwO}{}%
12. Fishler R, Hofemeier P, Etzion Y, Dubowski Y, Sznitman J. Particle dynamics and deposition in true-scale pulmonary acinar models. Sci Rep. 2015;5: 14071. doi:\href{https://doi.org/10.1038/srep14071}{10.1038/srep14071}

\leavevmode\hypertarget{ref-eb0gk6VO}{}%
13. Hasler D, Anagnostopoulou P, Nyilas S, Latzin P, Schittny J, Obrist D. A multi-scale model of gas transport in the lung to study heterogeneous lung ventilation during the multiple-breath washout test. Marsden AL, editor. PLoS Comput Biol. 2019;15: e1007079. doi:\href{https://doi.org/10.1371/journal.pcbi.1007079}{10.1371/journal.pcbi.1007079}

@@

\leavevmode\hypertarget{ref-CPhMy3Ie}{}%
18. Woodward JD, Maina JN. A 3D digital reconstruction of the components of the gas exchange tissue of the lung of the muscovy duck, Cairina moschata. J Anatomy. 2005;206: 477--492. doi:\href{https://doi.org/10.1111/j.1469-7580.2005.00413.x}{10.1111/j.1469-7580.2005.00413.x}

\leavevmode\hypertarget{ref-1GJECnKBu}{}%
19. Rodriguez M, Bur S, Favre A, Weibel ER. Pulmonary acinus: Geometry and morphometry of the peripheral airway system in rat and rabbit. Am J Anat. 1987;180: 143--155. doi:\href{https://doi.org/10.1002/aja.1001800204}{10.1002/aja.1001800204}

@@

\leavevmode\hypertarget{ref-RGBeCf8v}{}%
22. Kizhakke Puliyakote AS, Vasilescu DM, Newell JD, Wang G, Weibel ER, Hoffman EA. Morphometric differences between central vs. surface acini in A/J mice using high-resolution micro-computed tomography. J Appl Physiol. 2016;121: 115--122. doi:\href{https://doi.org/10.1152/japplphysiol.00317.2016}{10.1152/japplphysiol.00317.2016}

@@

\leavevmode\hypertarget{ref-knMAkPPz}{}%
27. Marone F, Hintermüller C, McDonald S, Abela R, Mikuljan G, Isenegger A, et al. X-ray Tomographic Microscopy at TOMCAT. J Phys: Conf Ser. 2009;186: 012042. doi:\href{https://doi.org/10.1088/1742-6596/186/1/012042}{10.1088/1742-6596/186/1/012042}

@@

\leavevmode\hypertarget{ref-VELl7OiR}{}%
29. Haberthür D, Hintermüller C, Marone F, Schittny JC, Stampanoni M. Radiation dose optimized lateral expansion of the field of view in synchrotron radiation X-ray tomographic microscopy. J Synchrotron Rad. 2010;17: 590--599. doi:\href{https://doi.org/10.1107/s0909049510019618}{10.1107/s0909049510019618}

\leavevmode\hypertarget{ref-dNc8FfNn}{}%
30. Hsia CCW, Hyde DM, Ochs M, Weibel ER. An Official Research Policy Statement of the American Thoracic Society/European Respiratory Society: Standards for Quantitative Assessment of Lung Structure. Am J Respir Crit Care Med. 2010;181: 394--418. doi:\href{https://doi.org/10.1164/rccm.200809-1522st}{10.1164/rccm.200809-1522st}

@@

\leavevmode\hypertarget{ref-nPoQ2EIB}{}%
33. Tschanz SA, Burri PH, Weibel ER. A simple tool for stereological assessment of digital images: the STEPanizer: TOOL FOR STEREOLOGICAL ASSESSMENT. Journal of Microscopy. 2011;243: 47--59. doi:\href{https://doi.org/10.1111/j.1365-2818.2010.03481.x}{10.1111/j.1365-2818.2010.03481.x}

\leavevmode\hypertarget{ref-QiAxY2i3}{}%
34. Hyde DM, Tyler NK, Putney LF, Singh P, Gundersen HJG. Total number and mean size of alveoli in mammalian lung estimated using fractionator sampling and unbiased estimates of the Euler characteristic of alveolar openings. Anat Rec. 2004;277A: 216--226. doi:\href{https://doi.org/10.1002/ar.a.20012}{10.1002/ar.a.20012}

@@

\leavevmode\hypertarget{ref-FE9HLB4f}{}%
36. Cruz-Orive LM. Precision of Cavalieri sections and slices with local errors. J Microsc. 1999;193: 182--198. doi:\href{https://doi.org/10.1046/j.1365-2818.1999.00460.x}{10.1046/j.1365-2818.1999.00460.x}

@@

\leavevmode\hypertarget{ref-IkHrgIj3}{}%
41. Shapiro SS, Wilk MB. An analysis of variance test for normality (complete samples). Biometrika. 1965;52: 591--611. doi:\href{https://doi.org/10.1093/biomet/52.3-4.591}{10.1093/biomet/52.3-4.591}

\leavevmode\hypertarget{ref-MyzxBJ57}{}%
42. Mann HB, Whitney DR. On a Test of Whether one of Two Random Variables is Stochastically Larger than the Other. Ann Math Statist. 1947;18: 50--60. doi:\href{https://doi.org/10.1214/aoms/1177730491}{10.1214/aoms/1177730491}

@@

\leavevmode\hypertarget{ref-8Miti2Gz}{}%
44. Virtanen P, Gommers R, Oliphant TE, Haberland M, Reddy T, Cournapeau D, et al. SciPy 1.0: fundamental algorithms for scientific computing in Python. Nat Methods. 2020;17: 261--272. doi:\href{https://doi.org/10.1038/s41592-019-0686-2}{10.1038/s41592-019-0686-2}

@@

\leavevmode\hypertarget{ref-I9TmP6IU}{}%
46. Zeltner TB, Bertacchini M, Messerli A, Burri PH. Morphometric Estimation of Regional Differences in the Rat Lung. Experimental Lung Research. 1990;16: 145--158. doi:\href{https://doi.org/10.3109/01902149009087879}{10.3109/01902149009087879}

\leavevmode\hypertarget{ref-Zv1XcgAO}{}%
47. Borisova E, Lovric G, Miettinen A, Fardin L, Bayat S, Larsson A, et al. Micrometer-resolution X-ray tomographic full-volume reconstruction of an intact post-mortem juvenile rat lung. Histochem Cell Biol. 2021;155: 215--226. doi:\href{https://doi.org/10.1007/s00418-020-01868-8}{10.1007/s00418-020-01868-8}

@@

\leavevmode\hypertarget{ref-eioib1TQ}{}%
49. Sznitman J, Sutter R, Altorfer D, Stampanoni M, Rösgen T, Schittny JC. Visualization of respiratory flows from 3D reconstructed alveolar airspaces using X-ray tomographic microscopy. J Vis. 2010;13: 337--345. doi:\href{https://doi.org/10.1007/s12650-010-0043-0}{10.1007/s12650-010-0043-0}

@@

\leavevmode\hypertarget{ref-YuJbg3zO}{}%
59. Himmelstein DS, Rubinetti V, Slochower DR, Hu D, Malladi VS, Greene CS, et al. Open collaborative writing with Manubot. Schneidman-Duhovny D, editor. PLoS Comput Biol. 2019;15: e1007128. doi:\href{https://doi.org/10.1371/journal.pcbi.1007128}{10.1371/journal.pcbi.1007128}
\end{cslreferences}
